# (PART) Linear Algebra in R {-}

# Class Examples

## Day 5: Tuesday May 25
```{r}
require('pracma')
```


Give the parametric solution to $A x = b$ if 
$$
A=\left[
\begin{array}{ccccc}
 1 & 1 & 1 & -1  \\
 1 & 0 & 3 & -1  \\
 -1 & 1 & -5 & 1  \\
 1 & 2 & -1 & 1  \\
\end{array}
\right] \qquad \hbox{and}\qquad  b = \begin{bmatrix} 2 \\ 4 \\ -6 \\ 4 \end{bmatrix}. \hskip.5in
$$
```{r,echo=TRUE}
A = cbind(c(1,1,-1,1),c(1,0,1,2),c(1,3,-5,-1),c(-1,-1,1,1))
b = c(2,4,-6,4)
Ab = cbind(A,b)
A
Ab
rref(Ab)
```
a. Give the solution this $Ax=b$ problem in *parametric* form:

$$
\begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \end{bmatrix}
= 
\begin{bmatrix}  6 \\ -2 \\ 0 \\ 2 \end{bmatrix}
+ s \begin{bmatrix}  -3 \\ 2 \\ 1 \\ 0 \end{bmatrix}
$$

b. Let's make two different solutions. One with s = 0 and one with s = 1. They are

```{r,echo=TRUE}
x1 = c(6,-2,0,2)
x2 = c(3,0,1,2)
```

And we check to see that they are solutions. Observe that we use the unfortunate notation `%*%` to multply a matrix times a vector, but also that we get b in both cases:
```{r,echo=TRUE}
A %*% x1
A %*% x2
```
Notice that we get b for both.

c. Now we look at the difference. It is a solution to the *homogeneous* equations A x = 0 as we see here:
```{r,echo=TRUE}
xh = x1 - x2
A %*% xh
```

d. Are any of these vectors solutions to Ax=b?

```{r,echo=TRUE}
A %*% (x1 + x2)
```
No! (this is 2b)

```{r,echo=TRUE}
A %*% (1/2*x1 + 1/2* x2)
```

Yes! This is b

```{r,echo=TRUE}
A %*% (x1 + 2021 * xh)
```
Yes! This is b.