
# Linear Transformations of a House


[Download this Rmd file from GitHub](https://github.com/Tom-Halverson/math236_s21/blob/main/04-house-transformations.Rmd)


Here is a plot of my house. You will need to run this chunk of code each time you re-start R to get the house back in memory.

```{r,house1,fig.height=4,fig.width=4, echo=TRUE}
house = cbind(c(0,0), c(0,3/4), c(1/2,3/4), c(1/2,0), c(1,0), c(1,1), c(5/4,1), c(0,2), c(-5/4,1), c(-1,1), c(-1,0), c(0,0));

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted")
polygon(house[1,], house[2,], col = "gray", border = "blue")
```

Here we explore linear transformations on the plane by looking at their effect on my house.


We give a series of examples of 2D linear transformations. After each example, it's your turn to play with variations from the same family of transformations. 

## Rotations

Suppose that we wish to *rotate* my house by pi/3 radians. As we've seen, a 2D rotation matrix by $t$ radians, counter-clockwise, is given by
$$A=\displaystyle{ \begin{bmatrix} \cos(t) & -\sin(t) \\ \sin(t) & \cos(t) \end{bmatrix}}.$$
Here is the code to display this transformation.
Observe that I apply the matrix `A` to the house, call it `house2` and plot both the original house and the new house in the same plot.

```{r,house2,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
t = pi/3
A = cbind(c(cos(t),sin(t)),c(-sin(t),cos(t)))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```

Try changing the angle above to achieve different rotations. How can you rotate it clockwise?


## Expansion and contraction
Next, we scale the house by 2 in the $x$-direction and by 3 in the $y$-direction.

```{r,house3,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
A = cbind(c(2,0),c(0,3))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```

Your turn: try some different scale factors. What if you use negative scale factors.


## Reflection

Now we reflect over the line y = x. 

```{r,house4,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
A = cbind(c(0,1),c(1,0))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```

Your turn: try the reflections (1)  over the x-axis; (2) over the y-axis; and (3) through the origin, i.e., sending (x,y) to (-x,-y).


## Shear Transformations

A shear transformation is of the form
$$
A=\displaystyle{ \begin{bmatrix} a & b \\ 0 & c \end{bmatrix}}
\quad
\mbox{and}
\quad
A=\displaystyle{ \begin{bmatrix} a & 0 \\ b & c \end{bmatrix}}
$$
For example:

```{r,house5,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
A = cbind(c(1,0),c(1,1))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```

You try: try to get the house to slant in the other direction.


## Dimension Reduction


Here we perform the transformation that sends  $\mathsf{e}_1$ to $(-1,1/2)$ and $\mathsf{e}_2$ to $(2,-1)$. Notice that they are the same line and the transformation projects the house onto this line.

```{r,house6,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
A = cbind(c(-1,1/2),c(2,-1))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```



## Your Turn

See if you can do the transformations in problem 3.3.

```{r,fig.height=4,fig.width=4, echo=TRUE}
# define the matrix A. This is the only part you should need to edit.
A = cbind(c(1,0),c(0,1))
A # display the matrix A

#----------------- this code applies the transformation and plots
# create a plot that we will add more layers to
house2 = A %*% house
plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # add  grid lines
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "pink", border = "black")
```



