

# Fractals 

```{r}
require(pracma)
```


Fractals are mathematical sets with a self-similar structure. Each smaller part is a miniature copy of the whole structure. Today, we will see how to use linear transformations and homogeneous coordintes to generate a fractal.


We will generate a fractal by repeatedlyapplying the following 3D linear transformations. 

```{r}
(A1 = cbind(c(.86,-.03,0),c(.03,.86,0),c(0,1.5,1)))
(A2 = cbind(c(.2,.21,0),c(-.25,.23,0),c(0,1.5,1)))
(A3 = cbind(c(-.15,.27,0),c(.25,.26,0),c(0,.45,1)))
(A4 = cbind(c(0,0,0),c(.0,.17,0),c(0,0,1)))
```
Observe that, via homogenous coordinates, each of them does some transformation in $\mathbb{R}^2$ and some translation.

We choose among these four transformations using some randomness: 83% of the time we use A1, 8% of the time we use A2, 8% of the time we use A3, and 1% of the time we use A4. This is done in the following function. 


```{r}
get_matrix <- function(x) {
  val = runif(1) # generate a random number between 0 an 1
  if (val < .83) {
    A1  # return A1
  } else if (val < .91) {
    A2  # return A2
  } else if (val < .99) {
    A3  # return A3 
  } else {
    A4   # return A4 
  }
}
```

Now, we start with the point (0,0,1) in homogeneous coordinates. And apply this function over and over again to the previous point that was just computed, plotting the new point each time. 

```{r,fig.height=5,fig.width=5}
pt = c(0,0,1) # starting point

plot(pt[1], pt[2], pch='.', xlim=c(-5,5), ylim=c(-1,12)) # set up the plot

for (i in 1:2000) # loop
  {
  points(pt[1], pt[2], pch='.',col='forest green')
  pt = get_matrix() %*% pt
}

```


## Transforming My House.

So what is happening? Let's look at the effect of each of these linear transformations on my house

Here is what A1 does to the house.  This happens 91% of the time.
```{r,fig.height=5,fig.width=5}

house = cbind(c(0,0,1), c(0,3/4,1), c(1/2,3/4,1), c(1/2,0,1), c(1,0,1), c(1,1,1), c(5/4,1,1), c(0,2,1), c(-5/4,1,1), c(-1,1,1), c(-1,0,1), c(0,0,1));

house2 = A1 %*% house

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted")
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "forest green", border = "black")
```

Here is what A2 does to the house. This happens 8% of the time.
```{r,fig.height=5,fig.width=5}

house = cbind(c(0,0,1), c(0,3/4,1), c(1/2,3/4,1), c(1/2,0,1), c(1,0,1), c(1,1,1), c(5/4,1,1), c(0,2,1), c(-5/4,1,1), c(-1,1,1), c(-1,0,1), c(0,0,1));

house2 = A2 %*% house

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted")
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "forest green", border = "black")
```
Here is what A3 does to the house. This happens 8% of the time.
```{r,fig.height=5,fig.width=5}

house = cbind(c(0,0,1), c(0,3/4,1), c(1/2,3/4,1), c(1/2,0,1), c(1,0,1), c(1,1,1), c(5/4,1,1), c(0,2,1), c(-5/4,1,1), c(-1,1,1), c(-1,0,1), c(0,0,1));

house2 = A3 %*% house

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted")
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "forest green", border = "black")
```
Here is what A4 does to the house. This happens 1% of the time.
```{r,fig.height=5,fig.width=5}

house = cbind(c(0,0,1), c(0,3/4,1), c(1/2,3/4,1), c(1/2,0,1), c(1,0,1), c(1,1,1), c(5/4,1,1), c(0,2,1), c(-5/4,1,1), c(-1,1,1), c(-1,0,1), c(0,0,1));

house2 = A4 %*% house

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-5,5),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted")
polygon(house[1,], house[2,], col = "gray", border = "blue")
polygon(house2[1,], house2[2,], col = "forest green", border = "forest green")
```
```{r,fig.height=5,fig.width=5}

house = cbind(c(0,0,1), c(0,3/4,1), c(1/2,3/4,1), c(1/2,0,1), c(1,0,1), c(1,1,1), c(5/4,1,1), c(0,2,1), c(-5/4,1,1), c(-1,1,1), c(-1,0,1), c(0,0,1));

plot(house[1,],house[2,],type="n",xlim=c(-5,5),ylim=c(-.5,10),xlab="x",ylab="y")
abline(h=-6:6, v=-6:6, col="gray", lty="dotted") # set up the plot

for (i in 1:1000) # loop
  {
  polygon(house[1,], house[2,], col = "forest green", border = "black")
  house = get_matrix() %*% house
}

```

##  Your Turn

Make some small adjustments to one of these matrices. Explore the impact of the fractal. Can you make it spikier?  Bushier?

```{r}
A1 = cbind(c(.86,-.03,0),c(.03,.86,0),c(0,1.5,1)) 
A2 = cbind(c(.2,.21,0),c(-.25,.23,0),c(0,1.5,1)) 
A3 = cbind(c(-.15,.25,0),c(.27,.27,0),c(0,.45,1)) 
A4 = cbind(c(0,0,0),c(.0,.17,0),c(0,0,1)) 
```

```{r,fig.height=5,fig.width=5}
pt = c(0,0,1)

plot(pt[1], pt[2], pch='.', xlim=c(-5,5), ylim=c(-1,12))
for (i in 1:1000) {
  points(pt[1], pt[2], pch='.')
  pt = get_lt() %*% pt
}

```


## Here are some from http://paulbourke.net/fractals/ifs/ 

             a     0.2020    0.1380
             b    -0.8050    0.6650
             c    -0.6890   -0.5020
             d    -0.3420   -0.2220
             e    -0.3730    0.6600
             f    -0.6530   -0.2770

```{r,fig.height=5,fig.width=5}

B1 = cbind(c(0.202,-.689,0),c(-.805,-0.342,0),c(-0.3730,-.653,1))
B2 = cbind(c(0.138,-.502,0),c(.665,-0.222,0),c(.66,-.2777,1))


get_bourke <- function() {
  val = rand()
  if (val < .6) {
    A = B1
  } else {
    A = B2
  }
  
  A
}


pt = c(0,0,1)

plot(pt[1], pt[2], pch='.', xlim=c(-.5,1), ylim=c(-1.25,.5))
for (i in 1:1000) {
  points(pt[1], pt[2], pch='.')
  pt = get_bourke() %*% pt
}
```

```{r}
get_list_ifs <- function(ifs) {
  vec <- list()
  for (i in 1:dim(ifs)[2]) {
    vec[[i]] = cbind(c(ifs[1,i],ifs[3,i],0),c(ifs[2,i],ifs[4,i],0),c(ifs[5,i],ifs[6,i],1))
  }
  vec
}

get_rand_idx <- function(p ) {
  ret_val  = 1
  q =  0
  x  = rand()
  for (i in 1:length(mat_list)) {
    q = q + p[i]
    if (x < q) {
      ret_val =  i
      break
    }
  }

  ret_val
}

bourke_dat = rbind(c(0.2020, 0.1380), c(-0.8050, 0.6650),
             c(-0.6890,-0.5020),
             c(-0.3420,-0.2220),
             c(-0.3730,0.6600),
             c(-0.6530,-0.2770))

bourke_list = get_list_ifs(bourke_dat)
bourke_p  =  c(.5,.5)


```

```{r}
dragon_list = rbind(c(0.824074,0.088272),
c( 0.281428,0.520988),
c(-0.212346,-0.463889),
c(0.864198,-0.377778),
c(-1.882290,0.785360),
c(-0.110607,8.095795))
dragon_p = c(0.8, 0.2)

dragon_list
```




```{r,fig.height=5,fig.width=5}
pt = c(0,0,1)
mat_list = get_list_ifs(dragon_list)
p = dragon_p

plot(pt[1], pt[2], pch='.', xlim=c(-7,7), ylim=c(0,12))

for (i in 1:1000) {
  points(pt[1], pt[2], pch='.')
  mat = mat_list[[get_rand_idx(p)]]
  pt = mat %*% pt
}
```
```{r,fig.height=3,fig.width=3}
spiral_data = rbind(c(0.787879,-0.121212, 0.181818),
c(-0.424242,0.257576,-0.136364),
c(0.242424,0.151515, 0.090909),
c(0.859848,0.053030, 0.181818),
c(1.758647,-6.721654, 6.086107),
c(1.408065,1.377236, 1.568035))



spiral_p = c(0.90,0.05, 0.05)

pt = c(0,0,1)
mat_list = get_list_ifs(spiral_data)
(mat_list)
p = spiral_p

plot(pt[1], pt[2], pch='.', xlim=c(-7,7), ylim=c(0,12))

for (i in 1:1000) {
  points(pt[1], pt[2], pch='.')
  mat = mat_list[[get_rand_idx(p)]]
  pt = mat %*% pt
}

```
