# Matrix Multiplication

Here we will practice multiplying matrices in R. First, let's define a few matrices.I'm using a trick here. By putting the assignment in parentheses, it assigns the matrix and displays it.s
```{r}
(A = cbind(c(1,2,3),c(4,5,6),c(1,1,-1)))
(B = cbind(c(1,-1,1),c(1,1,1),c(0,2,1)))
(C = cbind(c(2,1,1),c(1,0,1),c(1,-3,1),c(3,2,1)))
```
We multiply using the `%*%` command. As seen here:

```{r}
A %*% B
```

Note that 
```{r}
B %*% A
```
1. What do these last two multiplications say about the matrix product AB and BA? This is a very important property (or, perhaps, lack of property) of matrix multiplication.

2. Try multiplying BC and CB. What happens? And why?

```{r}

```

3. The *transpose* of a matrix is computed by `t(A)`. Compute the transpose of the matrices A, B, C and be sure that you all understand what it does. 

```{r}

```

4. The command `diag(n)` gives the n x n *identity matrix*. This is denoted $I_n$. For example, here is $I_3$. 
```{r}
diag(3)
```
Compute $I_2$, $I_4$, and $I_5$ and be sure you all agree on what the identity matrix is. 

5. Multiply the matrices A and B by the appropriately-sized identity matrix. Multiply  both ways, A I and I A, and agree upon what multiplying by the identity does.

```{r}

```

6. Multiply C by an identity matrix I C and C I. You might need a different size on the left and on the right.

```{r}

```

7. Our topic for Thursday (tomorrow) is the inverse of a matrix. You  compute the inverse of the matrix A  with `solve(A)`. Try this
```{r}
solve(B)
B %*% solve(B)
```

8. Multiply A by its inverse and look closely at the answer you get. 

```{r}

```

9. Some matrices do not have inverses. Try computing the inverse of the following matrices. We will discuss this tomorrow!

```{r}
(M1 = cbind(c(3,5),c(-2,1)))
(M2 = cbind(c(4,3),c(5,4)))
(M3 = cbind(c(4,2),c(10,5)))
```

10. Enter the matrix A in problem 3.7 in the homework. Then compute the matrix G which is A times its transpose. Discuss its meaning.

