
# Problem Set 4

* Due: Wednesday February 17 by noon CST. 
* Upload your solutions to Moodle in a PDF. 
* Please feel free to **use RStudio for all row reductions.**
* You can download the [Rmd source file  for this problem set](https://github.com/Tom-Halverson/math236_s21/blob/main/PS4-problem-set-4.Rmd).



The Problem Set covers sections 2.2-2.3 on Matrix Inverses and 4.1-4.3 on Subspaces and Bases.

```{r}
require('pracma')
```


## Rainy Day in LA

In Los Angeles if it rains today, there is a 50% chance it will rain tomorrow, but it if is sunny today, there is a 90% chance it will be sunny tomorrow. This is modeled in the rain-sunshine probability matrix P. 
$$
P = 
\begin{array}{c|cc|}
&\text{rain}&\text{sun}\\
\hline
\text{rain}&1/2&1/10\\
\text{sun}&1/2&9/10\\
\hline
\end{array}
$$
This matrix works as follows: if the rain-sunshine probability today is (40, 60) (that is, 40% chance rain and 60% chance sunshine), then the rain-sunshine probability tomorrow is (26, 74) as seen by the calculation below.
$$
\begin{bmatrix}
1/2 & 1/10 \\
1/2 & 9/10 \\
\end{bmatrix} \begin{bmatrix} 40 \\ 60 \end{bmatrix} = \begin{bmatrix} 26 \\ 74 \end{bmatrix}
$$

a. Find the rain-sunshine probability the day after tomorrow.

b. Compute $P^2$ and explain the meaning of each of the four entries in the matrix.

c. Find $P^{-1}$ and   and use it find the rain-sunshine probability *yesterday* if the rain-sunshine probability today is (40, 60).

## Fibonacci Vectors

The Fibonacci vectors $F$ in $\mathbb{R}^5$ are defined below:
$$
F  = \left\{ \begin{bmatrix} x_1 \\ x_2 \\ x_3 \\ x_4 \\ x_5 \end{bmatrix} ~\Bigg\vert~ \ x_3 = x_1 + x_2, x_4 = x_2 + x_3, x_5 = x_3 + x_4 \right\} \subseteq \mathbb{R}^5.   \hskip5in
$$

Find a basis for $F$. Be sure to show that your vectors span $F$ and are linearly independent.


## Vectors Rescaled

If the function $T: \mathbb{R}^n \to \mathbb{R}^n$ is a linear transformation, then show that the set below is a subspace of $\mathbb{R}^n$
$$
E_2 = \{\ \vec{x} \in \mathbb{R}^n \mid T(\vec{x}) = 2 \vec{x} \}.
$$
Is there anything special about 2 in the definition? If it were replaced by another scalar, would it still be a subspace?


## Column and Null Space

Find a basis for the column space $Col(A)$ and the null space $Nul(A)$ of the following matrix $A$ below
```{r,echo=TRUE}
(A = rbind(c(1, 2, 0, 2, 0, -1),c(1, 2, 1, 1, 0, -2),
           c(2, 4, -2, 6, 1, 2),c(1, 2,  0, 2, -1, -3 )))
```

## Extend to a basis

I am  interested in the vectors below. I know that they do not span $\mathbb{R}^5$, because there are not enough of them, but I want to extend this set to a basis of $\mathbb{R}^5$ by adding some vectors to the set. 
$$
\begin{bmatrix} 5\\ 4\\ 3\\ 1\\ 2 \end{bmatrix},
\begin{bmatrix} 4\\ 4\\ 3\\ 1\\ 2 \end{bmatrix}, 
\begin{bmatrix} 1\\ 1\\ 1\\ 1\\ 1\end{bmatrix}. 
$$
I searched online for ideas and one suggested that I make the matrix below and row reduce it.
```{r,echo=TRUE}
(A = cbind(c(5,4,3,1,2),c(4,4,3,1,2),c(1,1,1,1,1),diag(5)))
```

a. Row reduce this matrix.

b. Use the result to come up with a basis for $\mathbb{R}^5$ that includes my original 3 vectors

c. Explain why this method works.


## Getting Into a Subspace

Let $S \subset \mathbb{R}^n$ be a subspace and let $\mathsf{v}, \mathsf{w} \in \mathbb{R}^n$. For  each of the following statements, either give  a specific example  or explain why it cannot happen.


a. If $\mathsf{v}$ is in $S$ but  $\mathsf{w}$ is **not** in $S$, can  $\mathsf{v} + \mathsf{w}$ be in $S$?  



b. If $\mathsf{v}$ is **not** in $S$ and  $\mathsf{w}$ is **not** in $S$, can  $\mathsf{v} + \mathsf{w}$ be in $S$? 



c. If $\mathsf{v}$ is **not** in  $S$ and $c$ is a  nonzero constant, can $c\mathsf{v}$ be  in $S$?




## A Vector in Both Col(A) and Nul(A)

Give a $3 \times 3$ matrix $A$ for which the vector $\mathsf{v} = \begin{bmatrix}3 \\ -2 \\ 5   \end{bmatrix}$ is in both $\mathrm{Col}(A)$ and $\mathrm{Nul}(A)$. Be sure to demonstrate that $\mathsf{v} \in \mathrm{Col}(A)$ and $\mathsf{v} \in \mathrm{Nul}(A)$.


## Adjacency Matrix

You can do this problem in R or by hand. Consider the matrix $A$ defined here
```{r,echo=TRUE}
A = rbind(c( 0 , 1 , 0 , 1 , 1 ,0), c(1 , 0 , 1 , 1 , 0, 0 ),c( 0 , 1 , 0 , 1 , 1, 0 ),
  c( 1 , 1 , 1 , 0 , 1, 0 ),c( 1 , 0 , 1 , 1 , 0, 1 ), c(0, 0, 0, 0, 1, 0))
A
```

This matrix represents the connections in the network diagram below. There is a 1 in position $(i,j)$ of the matrix if there is a connection (an edge) between vertex $i$ and vertex $j$ and there is a 0 if there is not.

```{r,echo=FALSE}
suppressPackageStartupMessages(library(igraph))
g=graph_from_adjacency_matrix(A,mode='undirected')
V(g)$label = c(1,2,3,4,5,6)
plot(g)
```

a. Compute $A v$ where $v$ is the vector of all 1's. Explain what this new vector tells us about the graph.

b. Compute $A^2 = A A$, the square of the matrix $A$.

c. Look at the $(2,5)$ entry of $A^2$. Explain what this entry says about connections in the network. Do the same for the $(2,3)$ and the $(2,6)$ entry of $A^2$.

