---
title: "Untitled"
author: "Tom Halverson"
date: "1/29/2021"
output: html_document
---

title: "Problem 2.9"
author: "Your Name"
date: "02/02/2021"
output:
  html_document: default
  pdf_document: default
---

#### Enter the matrix A
We have to load the pracma package in order to get `rref`.
```{r}
require(pracma)
```

Here is the matrix A for this problem.
```{r}
A = cbind(c(12,-7,9,-4,8),c(10,-6,9,-3,7),c(-6,4,-9,-1,-5),c(8,-5,9,0,6),c(4,-7,9,-8,1),c(-18,16,-27,9,-12))
A
```
#### (a) Linear Dependence

Show that the columns of A are linearly independent by finding two different dependency relations among them. You can write your answer a form  like 5 a1+ 4 a2 + 3 a3 + 2 a4 + a5 = 0, where a1, a2, etc are the columns of A. 

```{r}
rref(A)
```

Using the $A x= 0$ solution $(-1,1,1,1,0,0)$ we get
$$
- \begin{bmatrix} 12 \\ -7 \\ 9 \\ -4 \\ 8 \end{bmatrix}
+ 
\begin{bmatrix} 10 \\ -6 \\ 9 \\ -3 \\ 7 \end{bmatrix}
+
\begin{bmatrix} -6 \\ 4 \\ -9 \\ -1 \\ -5 \end{bmatrix}
+
\begin{bmatrix}  8 \\  -5 \\ 9 \\ 0 \\ 6 \end{bmatrix}
= 
\begin{bmatrix} 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix}
$$
Using the $A x= 0$ solution $(-1,2,-1,0,1,1)$ we get
$$
- \begin{bmatrix} 12 \\ -7 \\ 9 \\ -4 \\ 8 \end{bmatrix}
+ 
2 \begin{bmatrix} 10 \\ -6 \\ 9 \\ -3 \\ 7 \end{bmatrix}
-
\begin{bmatrix} -6 \\ 4 \\ -9 \\ -1 \\ -5 \end{bmatrix}
+
\begin{bmatrix}  4 \\  -7 \\ 9 \\ -8 \\ 1 \end{bmatrix}
+
\begin{bmatrix} -18 \\ 16 \\ -27 \\ 9 \\ -12 \end{bmatrix}
= 
\begin{bmatrix} 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{bmatrix}
$$


#### (b) Ax = b

Augment A with b and show that A x = b is consistent and has infinitely many solutions.

```{r}
b = c(14,-12,9,-15,6)
Ab = cbind(A,b)
Ab
rref(Ab)
```
This system is consistent, and it has free variables, so there are infintely many solutions.

#### (c) Remove the redundancies

Remove the free-variable columns from A to get a new, smaller matrix A'. Show that A' x = b has a unique solution and say what that solution is.

```{r}
Ap = cbind(c(12,-7,9,-4,8),c(10,-6,9,-3,7),c(-6,4,-9,-1,-5),c(4,-7,9,-8,1))
Apb = cbind(Ap,b)
Apb
rref(Apb)
```

Has a unique solution. This is the one you get above by setting the free variables to 0.